International Relations are a Social Network:\\ A Latent Network Approach for Global Politics

### 2017-10-21 paper/4_app.tex
more clean up.
still needed:

new:
reiter and stam graphics
mcdonald graphics

We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.  Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.

Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.

The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies that use a framework that ignores dependencies.  We believe that it does, and that the dynamic latent factor model provides a step forward.

@@

@@

\begin{figure}
	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
\end{figure}

\begin{figure}
\centering   
	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
\end{figure}

@@

THIS TABLE WILL BECOME A FIGURE, and skedaddle.

\begin{table}
\begin{center}
\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
\begin{tabular}{lrr} \toprule
& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
Variable & \texttt{logit} & \texttt{amen} \\ \midrule
Allied & 0.142&  0.023 \\
Joint Democracy &\bf -0.507&  0.045 \\
Peace Years \bf &-0.260& \bf -0.060 \\
Spline 1 &\bf -0.001& \bf 0.00 \\
Spline 2&\bf -0.000& \bf 0.00\\
Spline 3 &-0.000& \bf 0.00\\ 
Contiguity &\bf 2.412& \bf 0.640 \\
Parity &0.075&  -0.013 \\
Parity at entry year&\bf 0.868&  0.002 \\ 
Rivalry &\bf 2.031&\bf 0.721\\ 
Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
\end{tabular}\\
\end{center}
{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017). For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation. NEED TO BE FIXED if staying.
\end{table}

We re-estimated model 6 from Table 6 (2017, 34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).  Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.  Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$). In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.  As might be expected, the strong geographic clustering in the original study is about one-quarter as strong in the AME estimations. Similarly, rivalry coefficients are about one-third the size in the AME formulation, but a great deal more precisely measured ($z=18.116$).

We utilized the original and the AME results from Gibler's model 6 in Table 6 (as above) to calculate the expected values for one scenario. We focused on the variables measuring rivalry.  We used mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.  The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.  As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ from the GLM in two distinct ways. First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably higher when taking interdependencies into account with the AME model.  These are rare events, so the probabilities are low, but the difference is a factor of $3x$. Thus, you get quite substantially different expected values from these two models.  Second, the GLM model has a standard errors which underestimate the true variability, in part because some of them are assumed to be zero.  This shows up in the narrow density of predicted outcomes for the GLM model, compared to the density seen with the AME. The latter reflects a wider range of potential outcomes, reflecting the incorporation of uncertainty in the AME model.

\begin{figure}
	\caption{}
	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
 	\label{fig:gibmargeff}
 \end{figure}

>>>>

We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.  Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.  
Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.

The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies that use a framework that ignores dependencies.  Based in part our our results as well as the statistical and mathematical characteristics of the
AME framework, we believe that it does. The dynamic latent factor model provides a step forward in the modeling of international relations.

@@

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
\end{figure}
\FloatBarrier

@@

\begin{figure}
\centering   
	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
\end{figure}
\FloatBarrier

@@

Insert coefplot here, without splines...\label{fig:giberrepl}
%
%\begin{table}
%\begin{center}
%\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
%\begin{tabular}{lrr} \toprule
%& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
%Variable & \texttt{logit} & \texttt{amen} \\ \midrule
%Allied & 0.142&  0.023 \\
%Joint Democracy &\bf -0.507&  0.045 \\
%Peace Years \bf &-0.260& \bf -0.060 \\
%Spline 1 &\bf -0.001& \bf 0.00 \\
%Spline 2&\bf -0.000& \bf 0.00\\
%Spline 3 &-0.000& \bf 0.00\\ 
%Contiguity &\bf 2.412& \bf 0.640 \\
%Parity &0.075&  -0.013 \\
%Parity at entry year&\bf 0.868&  0.002 \\ 
%Rivalry &\bf 2.031&\bf 0.721\\ 
%Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
%\end{tabular}\\
%\end{center}
%{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017). For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation. NEED TO BE FIXED if staying.
%\end{table}

We re-estimated model 6 from Table 6 (2017, 34). The results are presented in Figure~\ref{fig:giberrepl} in order to facilitate explicit comparison.
The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).  Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.  Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$). In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.  As might be expected, the strong geographic clustering in the original study is about one-quarter as strong in the AME estimations. Similarly, rivalry coefficients are about one-third the size in the AME formulation, but a great deal more precisely measured ($z=18.116$).

We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario. We focused on the variables measuring rivalry.  We used mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.  The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.  As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ from the GLM in two distinct ways. First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably higher when taking interdependencies into account with the AME model.  These are rare events, so the probabilities are low, but the difference is a factor of $3x$. Thus, you get quite substantially different expected values from these two models.  Second, the GLM model has a standard errors which underestimate the true variability, in part because some of them are assumed to be zero.  This shows up in the narrow density of predicted outcomes for the GLM model, compared to the density seen with the AME. The latter reflects a wider range of potential outcomes, reflecting the incorporation of uncertainty in the AME model.

\begin{figure}
	\caption{THIS NEEDS TO BE FIXED.  \label{fig:gibmargeff}}
	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
 	\label{fig:gibmargeff}
 \end{figure}
